Preparing data from the Global Burden of Disease for METAHIT modelling

### 2021-10-01 data-raw/gbd_process.Rmd
Major update, featuring fast estimation method

@@

gbddb <- gbdmod %>%
  filter(measure %in% c("Deaths","Incidence", "Prevalence")) %>%
  mutate(measure = fct_recode(measure, mort="Deaths", inc="Incidence", prev="Prevalence")) %>% 
  pivot_wider(names_from=measure,
              values_from=c("num","denom"),
              values_fill = list(num=0, denom=5000)) %>% 
  mutate(location = fct_recode(str_to_title(location),
                                 `Greater Manchester`="Greatermanchester",
                                 `North East (city region)`="Northeast",
                                 `North East (region)`="North East",
                                 `West Midlands (city region)`="Westmidlands",
                                 `West Midlands (region)`="West Midlands",
                                 `East of England`="East Of England")) %>%
  mutate(cause = str_to_sentence(cause)) %>%
  arrange(location, cause, sex, ageyr) %>%
  filter(! cause %in% c("Hypertensive heart disease"))

@@

```{r}
library(readxl)
rem <- read_excel("~/work/chronic/Figure_10__Predicted_10-year_net-survival_using_the_hybrid_approach_for_men_and_women_(aged_15_to_99_years).xls", range="A7:C26", na=":") %>%
  pivot_longer(cols=c("Men","Women"), names_to = "sex", values_to = "surv10") %>% 
  mutate(prem1 = 1 - (1 - surv10/100)^(1/10)) %>% 
  filter(Cancer %in% c("Breast","Uterus","Colorectal","Liver","Stomach","Lung"),
         !is.na(prem1)) %>%
  mutate(Cancer = fct_recode(Cancer, 
                             "Breast cancer"="Breast",
                             "Uterine cancer"="Uterus",
                             "Colon and rectum cancer"="Colorectal",
                             "Stomach cancer"="Stomach",
                             "Tracheal, bronchus, and lung cancer"="Lung"
                             #,
                             #"Liver cancer"="Liver"
                             ),
         sex = fct_recode(sex, "Male"="Men", "Female"="Women")) %>% 
  rename(cause = Cancer)

@@

num_cancer <- gbddb %>% 
  filter(cause %in% cancers, 
         ageyr %in% 15:99) %>% 
  group_by(sex, cause) %>% 
  summarise(ncases = sum(num_prev)) %>% 
  left_join(rem, by=c("cause","sex")) %>%
  mutate(num_rem = round(ncases*prem1 / 100),
         denom_rem = round(ncases / 100)) %>% 
  select(sex, cause, num_rem, denom_rem)
num_cancer

gbddb <- gbddb %>% 
  left_join(num_cancer) %>% 
  mutate(num_rem = replace_na(num_rem, 0),
         denom_rem = replace_na(denom_rem, 0))

saveRDS(gbddb, file=file.path(relative_path_gbd, "gbddb.rds"))
saveRDS(gbddb, file=file.path("~/work/chronic", "gbddb.rds"))

>>>>

Rename variables for consistency with disbayes

@@

gbddb <- gbdmod %>%
  filter(measure %in% c("Deaths","Incidence", "Prevalence")) %>%
  mutate(measure = fct_recode(measure, mort="Deaths", inc="Incidence", prev="Prevalence")) %>% 
  pivot_wider(names_from=measure,
              values_from=c("num","denom"),
              values_fill = list(num=0, denom=5000)) %>% 
  mutate(area = fct_recode(str_to_title(location),
                           `Greater Manchester`="Greatermanchester",
                           `North East (city region)`="Northeast",
                           `North East (region)`="North East",
                           `West Midlands (city region)`="Westmidlands",
                           `West Midlands (region)`="West Midlands",
                           `East of England`="East Of England")) %>%
  mutate(disease = str_to_sentence(cause)) %>%
  rename(age = ageyr,
         gender = sex,
         inc_num = num_inc, inc_denom = denom_inc,
         prev_num = num_prev, prev_denom = denom_prev,
         mort_num = num_mort, mort_denom = denom_mort
  ) %>% 
  select(age, gender, disease, area, areatype, inc_num, inc_denom, prev_num, prev_denom, mort_num, mort_denom) %>%
  arrange(area, disease, gender, age) %>%
  filter(! disease %in% c("Hypertensive heart disease"))

@@

```{r}
library(readxl)
rem <- read_excel("~/work/chronic/cancerdata/Figure_10__Predicted_10-year_net-survival_using_the_hybrid_approach_for_men_and_women_(aged_15_to_99_years).xls", range="A7:C26", na=":") %>%
  pivot_longer(cols=c("Men","Women"), names_to = "sex", values_to = "surv10") %>% 
  mutate(prem1 = 1 - (1 - surv10/100)^(1/10)) %>% 
  filter(Cancer %in% c("Breast","Uterus","Colorectal","Liver","Stomach","Lung"),
         !is.na(prem1)) %>%
  mutate(Cancer = fct_recode(Cancer, 
                             "Breast cancer"="Breast",
                             "Uterine cancer"="Uterus",
                             "Colon and rectum cancer"="Colorectal",
                             "Stomach cancer"="Stomach",
                             "Tracheal, bronchus, and lung cancer"="Lung"
                             #,
                             #"Liver cancer"="Liver"
                             ),
         gender = fct_recode(sex, "Male"="Men", "Female"="Women")) %>% 
  rename(disease = Cancer)

@@

num_cancer <- gbddb %>% 
  filter(disease %in% cancers, 
         age %in% 15:99) %>% 
  group_by(gender, disease) %>% 
  summarise(ncases = sum(prev_num)) %>% 
  left_join(rem, by=c("disease","gender")) %>%
  mutate(rem_num = round(ncases*prem1 / 100),
         rem_denom = round(ncases / 100)) %>% 
  select(gender, disease, rem_num, rem_denom)
num_cancer

gbddb <- gbddb %>% 
  left_join(num_cancer) %>% 
  mutate(rem_num = replace_na(rem_num, 0),
         rem_denom = replace_na(rem_denom, 0))

saveRDS(gbddb, file=file.path(relative_path_gbd, "gbddb.rds"))
saveRDS(gbddb, file=file.path("~/work/chronic/disbayes/metahit", "gbddb.rds"))
